### 2015-06-05 lp.tex
address some comments

%!TEX root =  autocontgrlp.tex
\section{Approximate Linear Programming}
The LP formulation in \eqref{mdplp} can be represented in short as,
\begin{align}
\begin{split}
\label{mdplpshort}
\min_{J\in \R^n}\, &c^\top J\\
\text{s.t}\mb &J\geq T J,
\end{split}
\end{align}
or 
%\text{ or }\nn\\
\begin{align}
\begin{split}
\label{shortalter}
\min_{J\in \R^n} \,&c^\top J\\
\text{s.t}\mb &EJ\geq H J,
\end{split}
\end{align}
where $J\geq TJ$ is a shorthand for the $nd$ constraints in \eqref{mdplp} and $E$ is an $nd\times n$ matrix given by $E=[I,\ldots,I]^\top$, i.e., $E$ is obtained by stacking $d$ identical $n\times n$ identity matrices one over the other. Note that \eqref{mdplpshort} and \eqref{shortalter} are identical programs and differ only in notation. We use notation of type \eqref{mdplpshort} whenever we prefer brevity and we use notation \eqref{shortalter} in some definitions and proof for the sake of clarity.
The approximate linear program (ALP) is obtained by making use of LFA in the LP, i.e., by letting $J=\Phi r$ in \eqref{mdplpshort} and is given as
\begin{align}\label{alp}
\begin{split}
\min_{r\in \R^k}\, &c^\top \Phi r\\
\text{s.t}\mb &\Phi r\geq T \Phi r.
\end{split}
\end{align}
Unless specified otherwise we use $\tr$ to denote any solution\todoc{instead of saying the solution} to the ALP and $\tj=\Phi \tr$ to denote the corresponding approximate value function. We now state the assumptions and definitions used for the rest of the paper.
\begin{assumption}\label{one}
The first column of the feature matrix $\Phi$ (i.e., $\phi_1$) is $\one \in \R^n$. In other words, the constant function is part of the basis.
\end{assumption}
\begin{assumption}\label{probdist}
$c=(c(i),i=1,\ldots,n)\in \R^n$ is a probability distribution, i.e., $c(i)\geq 0$ and $\sum_{i=1}^n c(i)=1$.
\end{assumption}
\begin{definition}
Given a function $\chi\colon S\ra \R^+$ we define the quantity $\beta_{\chi}$ as
\begin{align*}
\beta_{\chi}=\max_{s \in S} \frac{\underset{a \in A}{\max}\big(\alpha\sum_{s'}p_a(s,s')\chi(s')\big)}{\chi(s)}\,.
\end{align*}
\end{definition}
\begin{definition}
The function $\chi$ is then said to be a \emph{Lyapunov} function if $\beta_{\chi}<1$.
\end{definition}
\begin{assumption}\label{lyap}
$\psi\colon S \ra \R^+$ is a Lyapunov function and is present in the column span of the feature matrix $\Phi$.
\end{assumption}
It is straightforward to check that the function $\one$ is a Lyapunov function and trivially is present in the column span of $\Phi$. 
\begin{definition}\label{modlone}
The weighted $L_1$-norm is defined as \todoc{modified $\Rightarrow$ weighted?}
\begin{align*}
||J||_{1,c}=\sum_{s \in S} c(s)|J(s)|,
\end{align*}
where $c$ obeys Assumption~\ref{probdist}.
\end{definition}
\begin{definition}\label{modnorm}
The weighted $L_\infty$-norm is defined as follows:  \todoc{modified $\Rightarrow$ weighted? Also, the convention in math is to divide in the case of the $L_{\infty}$ norm. Not a high priority to change this, but maybe it would be good to do this in the longer term.}
\begin{align*}
||J||_{\infty,\rho}=\max_{s \in S} \frac{|J(s)|}{\rho(s)},
\end{align*}
where $\rho \colon \S \ra \R^+$ is the weighing function and $J\in \R^n$.
\end{definition}
In Definition~\ref{modnorm}, the use of the weighting function $\rho$ allows us to measure the error taking into account the relative importance of the various states. A lower value of $\rho(s)$ means that the state $s$ is less important and vice-versa.\\
We recall below Theorem~$4.2$ of \cite{ALP} which bounds the error in the approximate value function.
\begin{theorem}\label{restateval}
Let $\tr$ be the solution to the ALP in \eqref{alp}, $\tilde{J}_c=\Phi \tilde{r}_c$, $\psi$ be a Lyapunov function and $c$ be a distribution as in \eqref{probdist}, then 
\begin{align*}
||J^*-\tj||_{1,c}\leq \frac{2c^\top \psi}{1-\beta_{\psi}}\min_{r}||J^*-\Phi r||_{\mn}.
\end{align*}
\end{theorem}
We now recall Theorem~$3.1$ of \cite{ALP} that characterizes the loss in performance of the greedy policy.
\begin{theorem}\label{restatepol}
Let $\tilde{u}$ be the greedy policy with respect to the solution $\tj$ of the ALP, then 
\begin{align*}
||J^*-J_{\tu}||_{1,c}\leq \frac{1}{1-\alpha}||J^*-\tj||_{1,c'},
\end{align*}
where $c'=(1-\alpha)c^\top(I-\alpha P_{\tu})^{-1}$.
\end{theorem}
Theorems~\ref{restateval} and ~\ref{restatepol} together imply that the ALP addresses both the control and prediction problems. Please refer to \cite{ALP} for a more detailed treatment of the ALP.\\
Note that the ALP is a linear program in $k$ ($<<n$) variables as opposed to the LP in \eqref{mdplpshort} which has $n$ variables. Nevertheless, the ALP has $nd$ constraints (same as the LP) which is an issue when $n$ is large and calls for constraint approximation/reduction techniques.