### 2016-07-08 single/autocontgrlp.tex

\documentclass[12pt,draftcls,onecolumn]{IEEEtran}
%\documentclass[12pt,twocolumn]{IEEEtran}
\usepackage{times}
\input{pack.tex}
%\documentstyle[nips14submit_09,times,art10]{article} % For LaTeX 2.09
\title{A Generalized Reduced Linear Program for Markov Decision Processes}
\author{Chandrashekar Lakshminarayanan and Shalabh Bhatnagar\\Department of Computer Science and Automation\\ Indian Institute of Science\\ Bangalore-560012\\\{chandrul,shalabh\}@csa.iisc.ernet.in\\ and \\
Csaba Szepesv\'{a}ri\\Department of Computing Science\\University of Alberta\\Edmonton, Alberta\\Canada T6G 2E8\\  csaba.szepesvari@ualberta.ca }

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.
\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}
%\nipsfinalcopy % Uncomment for camera-ready version
\begin{document}
\maketitle
\input{abstract}
%\input{intro}
\input{introshort}
\input{mdpshort}
\input{lp}
%\input{constsamp}
\input{grlp}
\input{errana}
%\input{lubp}
%\input{alubp}
%\input{strel}
\input{improv}
\input{polbnd}
\input{resdisc}
\input{num}
\input{concl}
%\input{example}
%\input{discussion}
\bibliographystyle{plain}
\bibliography{ref.bib}
\newpage
\onecolumn
%\input{appendix}
\end{document}