### 2016-07-16 appendix.tex
describing N and fixing a bug

\begin{appendix}
\textbf{On choice of $\N$:}\\
Let $g_{\min}=\underset{a,s}{\min}g_a(s)$, $g_{\max}=\underset{a,s}{\max}g_a(s)$ and define $r^*\eqdef\underset{r\in \R^k}{\arg\min}||J^*-\Phi r||_{\mn}$. Now $\frac{g_{\min}}{1-\alpha}\leq J^* \leq \frac{g_{\max}}{1-\alpha}$ and since $\mathbf{1}$ is in the basis, we have $||J^*-\Phi r^*||_{\mn}\leq \frac{g_{\max}-g_{\min}}{1-\alpha}$ (making use of the fact that for any sensible choice of Lyapunov function $||J^*-\Phi r^*||_{\mn}\leq ||J^*-\Phi r^*||_{\infty}$). Further, in our arguments, we have made the implicit assumption that $r'= r^*+||J^*-\Phi r^*||_{\mn} \left(\frac{1+\beta_{\psi}}{1-\beta_{\psi}}\right)\, r_0 \in \N$. This implicit assumption will continue to hold when $\N\eqdef\{\r| a\leq \Phi r\leq b\}$, where $a=\frac{g_{\min}}{1-\alpha}$ and $b=\frac{g_{\max}}{1-\alpha}+\frac{1+\beta_{\psi}}{1-\beta_{\psi}}\frac{g_{\max}-g_{\min}}{1-\alpha}$.
\end{appendix}