### 2017-03-20 talk-abstract.tex

Title: Towards a rigorous approach to approximate computations in huge-scale Markovian decision processes
Based on joint with Chandrashekar Lakshminarayanan and Shalabh Bhatnagar.

In this talk, without assuming any background beyond familiarity with basic probability and 
linear algebra, I will describe exciting new results on the computation of near-optimal policies in 
huge-scale Markovian decision processes (MDPs) via the so-called approximate linear programming methodology.
In a nutshell, the new results provide meaningful upper bounds on the quality of the computed policies
as a function of the linear basis functions chosen by the algorithm designer, while avoiding unrealistic assumptions
which have effectively become the norm in prior results. 
As opposed to previous work that relied on worst-case reasoning over randomly sampled constraints,
in this work the analysis is done using an operator-theoretic approach.
I will finish by discussing the remaining main open problems, connections to alternative approaches,
as well as potential extensions.
Markov decision processes lie at the hear of many reinforcement learning algorithms and have found
numerous applications across many engineering and scientific problems. 
----
Sequential decision making under uncertainty is a common element of numerous important scientific and engineering problems.
These problems are often phrased as Markov decision processes
Many important scientific and engineering problems involve sequential decision making under uncertainty.
----
MDPs are often used to model the dilemma of an agent who is trying to control its stochastic environment,
has widespread applications in science and engineering and is at the heart of formulating reinforcement learning problems.
Markovian decision processes merge Markov chains with elements of control:
The controlling agent observe a state, chooses an action, which results in a stochastic Markov transition 
to a next state while incurring some reward. The goal is to maximize the total expected discount reward.
While many problems can be modeled as MDPs, the number of states in these MDPs is more often than not is
at a scale so that only sublinear computations in the number of states are feasible. 
A common methodology uses compact representations based on linear function approximation.
The question then becomes how to interact with an oracle that gives a local, state-based access to the MDP
to efficiently compute a near-optimal policy.
----

to

Since Bellman it has been widely recognized that value-functions, mapping states of the process to reals, representing
the "value" of being in a given state, lie at the heart

A common computation model assumes state-based oracle access to the MDP
The algorithms that compute actions of a near-optimal "policy" can obtain averages of any selec

tend to have a state  where the number of states is intractably large,
A key problem is to compute given a large-scale model
Without assuming background knowledge on MDPs and reinforcement learning,

in this talk 
describe exciting new results on using approximate linear programming

building on linear programming

A common element of numerous important scientific and engineering problems is that

MDPs math model to capture seq. dec. making.
A discounted MDP is composed of states, actions, next-state distributions for each state-action pair, a reward
for each state-action pair and a discount factor in the (0,1) open interval.
An agent in an MDP, knowing which state it is at,
takes an action to arrive stochastically at some next state while incurring some reward.
An optimal behavior is one that maximizes the total expected discounted reward from any state.
The computational question is

The goal is to compute a policy, a map from states to actions, that maximizes from each state the total
expected discounted reward over time.
At a high level one is given finitely many Markov transition matrices, each matrix corresponding to one action.
The matrices capture stochastic transitions under the effect of an action.

If MDPs are the math model, how do we compute good
Core

What is the problem?
Computation of near-optimal policies in large scale MDPs.
State-based oracle access to large scale MDP, features for capturing the optimal value function.

Why should we care?

What's left?
Is this even asking the right question? Search in policy space vs. value space (direct vs. indirect).

This is a non-adaptive method.
Would adaptive methods work better?

Linear vs. nonlinear.

Applications of the LP approach to ADP range from scheduling in queueing networks (Morrison and Kumar., 1999; Veatch, 2005; Moallemi et al., 2008), revenue management (Adelman, 2007; Farias and Van Roy, 2007; Zhang and Adelman, 2008), portfolio management (Han, 2005), inventory problems (Adelman, 2004; Adelman and Klabjan, 2009), and algorithms for solving stochastic games (Farias et al., 2008) among others.

Approximate linear programming (ALP) and its variants have been widely applied to Markov Decision Processes (MDPs) with a large number of states. A serious limitation of ALP is that it has an intractable number of constraints, as a result of which constraint approximations are of interest. In this paper, we define a generalized reduced linear program (GRLP) that has a tractable number of constraints, which are obtained as positive linear combinations of the original constraints of the ALP. The main contribution of this paper is a novel theoretical framework developed to obtain error bounds for any given GRLP. By providing a detailed error analysis for the GRLP, we justify usage of a linear architecture for approximating the dual variables. Unlike prior results on constraint sampling, our analysis is deterministic and is based on a novel contraction operator.
\end{abstract}
\begin{keywords}{
Approximate Dynamic Programming (ADP), Markov Decision Processes (MDPs), Approximate Linear Programming (ALP), Generalized Reduced Linear Program (GRLP), Constraint Sampling, Reinforcement Learning.}
\end{keywords}