### 2017-03-23 appendix.tex
towards main result

\begin{appendix}
\textbf{On choice of $\N$:}
It is desirable to get rid of the assumption that $\tr\in \N$ without affecting our arguments. Note that the set $\N$ is not used in ALP \eqref{alp} and the feasible set of ALP can be specified as the set $\S\eqdef\{r\in \Re^k| \Phi r \geq T\Phi r\}$. Further, the only property about $\S$ that we used in our argument (\Cref{restate}) was that $r'= r^*+\norm{J^*-\Phi r^*}_{\mn} \left(\frac{1+\beta_{\psi}}{1-\beta_{\psi}}\right)\, r_0 \in \S$. Thus, so long as $\N$ contains $r'$ it is not necessary for it to contain $\tr$.\par
Let $g_{\min}=\underset{a,s}{\min}g_a(s)$, $g_{\max}=\underset{a,s}{\max}g_a(s)$ and define $r^*\eqdef\underset{r\in \Re^k}{\arg\min}\norm{J^*-\Phi r}_{\mn}$. Now $\frac{g_{\min}}{1-\alpha}\leq J^* \leq \frac{g_{\max}}{1-\alpha}$ and since $\mathbf{1}$ is in the basis, we have $\norm{J^*-\Phi r^*}_{\mn}\leq \frac{g_{\max}-g_{\min}}{1-\alpha}$ (making use of the fact that for any sensible choice of Lyapunov function $\norm{J^*-\Phi r^*}_{\mn}\leq \norm{J^*-\Phi r^*}_{\infty}$).  The fact that $r'\in \N$ can then be ensured when $\N\eqdef\{\r| a\leq \Phi r\leq b\}$, where $a=\frac{g_{\min}}{1-\alpha}$ and $b=\frac{g_{\max}}{1-\alpha}+\frac{1+\beta_{\psi}}{1-\beta_{\psi}}\frac{g_{\max}-g_{\min}}{1-\alpha}$.
\end{appendix}