### 2017-03-23 cone.tex
editing cone file

%!TEX root =  autocontgrlp.tex
\begin{theorem}
Let $\phi(s)$%
\footnote{Here $\phi(s)$ is the $s^{th}$ row of the feature matrix $\Phi$} be the feature of the $s^{th}$ state. Let $i$ be a state whose constraints have not been retained and let $\{J\}$ denote the set of states whose constraints have been retained. Further, let $\phi(i)\in \Re^k$ lie in the linear span of vectors $\phi(j)\in \Re^k,\forall j\in \J$, then
\begin{align*}
\MoveEqLeft (\Gamma J^*-\hg J^*)(i)
 \leq 
(\Gamma J^*-J^*)(i) \\
& {}+J^*(i)-\underset{j\in \J}{\sum} |\gamma_j| J^*(j)
+ \underset{j\in \J}{\sum}(|\gamma_j|-\gamma_j) B,
\end{align*}
where $\gamma_j$ are the co-efficients of the linear span and $B\eqdef \underset{\underset{s\in S}{r\in \N}}{\max}(\Phi r) (s)$
\end{theorem}
\begin{proof}
Let $r^s_{\Gamma}$ and $r^s_{\hg}$ be such that $(\Phi r^s_{\Gamma})(s)=(\Gamma J^*)(s)$ and $(\Phi r^s_{\hg})(s)=(\hg J^*)(s)$
\begin{align*}
\MoveEqLeft (\Gamma J^*-\hg J^*)(i) 
=\phi(i) r^s_{\Gamma}- \phi(i) r^s_{\Gamma}\\
&=\phi(i) r^i_{\Gamma}- \underset{j\in \J}{\sum} (|\gamma_j|+\gamma_j -|\gamma_j|)\phi(j) r^i_{\hg}\\
&\leq \phi(i) r^i_{\Gamma}- \underset{j\in \J}{\sum} |\gamma_j| J^*(j) + \underset{j\in \J} {\sum} (|\gamma_j|-\gamma_j)B\\
&= \phi(i) r^i_{\Gamma}- J^*(i)+ J^*(i) -\underset{j\in \J}{\sum} |\gamma_j| J^*(j) \\
& \quad + \underset{j\in \J}{\sum} (|\gamma_j|-\gamma_j)B\,.
\end{align*}
\end{proof}