### 2017-03-24 plt.tex

\FloatBarrier
\begin{figure}[htp]
\begin{minipage}{0.5\textwidth}
\resizebox{1.0\textwidth}{!}{
\begin{tabular}{cc}
\begin{tikzpicture}[scale=1,font=\Large,]
    \begin{axis}[]
    \addplot[smooth,very thick, mark=diamond, each nth point=100] plot file {./lralp_exp/V_LP};
    \addplot[dashed,very thick, mark=+, each nth point=100] plot file {./lralp_exp/V_pol_CS};
    \addplot[dotted,very thick,mark=square, each nth point=100] plot file {./lralp_exp/V_pol_cone};
    \addlegendentry{$J^*$};
    \addlegendentry{$J_{CS}$};
    \addlegendentry{$J_{LRA}$};

\end{axis}
\end{tikzpicture}

&
\begin{tikzpicture}[scale=1,font=\Large]
    \begin{axis}[]
\addplot[smooth,very thick] plot file {./lralp_exp/policy};
\addplot[dashed,very thick] plot file {./lralp_exp/policy_CS};
\addplot[dotted,very thick] plot file {./lralp_exp/policy_cone};
\addlegendentry{$u^*$};
\addlegendentry{$u_{CS}$};
\addlegendentry{$u_{LRA}$};

    \end{axis}
    \end{tikzpicture}

\end{tabular}
}
\end{minipage}
\end{figure}